{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Models for detecting BHBH, BHNS and NSNS binaries with LISA

### 2021-11-19 paper/tex/sections/comparisons.tex
Change Figure->Fig.

In Figure~\ref{fig:previous_studies}, we compare our results to similar previous studies that investigate the population of stellar-mass BHBHs, BHNSs and NSNSs that are detectable with LISA. Figure~\ref{fig:previous_studies} details the expected detection rates predicted by each paper as well as their assumptions regarding their Milky Way galaxy model, binary population synthesis simulation and LISA mission specifications. We only include papers that are similar to our work, such that they use population synthesis and simulate sources in the Galactic plane. Moreover, Figure~\ref{fig:previous_studies} does not include the numerous papers on the LISA WDWD population as we do not make predictions for these DCOs.

>>>>

In Fig.~\ref{fig:previous_studies}, we compare our results to similar previous studies that investigate the population of stellar-mass BHBHs, BHNSs and NSNSs that are detectable with LISA. Fig.~\ref{fig:previous_studies} details the expected detection rates predicted by each paper as well as their assumptions regarding their Milky Way galaxy model, binary population synthesis simulation and LISA mission specifications. We only include papers that are similar to our work, such that they use population synthesis and simulate sources in the Galactic plane. Moreover, Fig.~\ref{fig:previous_studies} does not include the numerous papers on the LISA WDWD population as we do not make predictions for these DCOs.

### 2021-11-19 paper/tex/sections/results.tex
Change Figure->Fig.

\subsection{Properties of the detectable systems}\label{sec:fiducial_distributions}
In Figure~\ref{fig:fiducial_pdf_distributions}, we show the 1D distribution of several individual parameters of the population of detectable binaries together with the 1- and 2-$\sigma$ uncertainties obtained via bootstrapping. These uncertainties represent the fluctuations in our results over different random instances of the Milky Way. The distributions shown here are approximated by kernel density estimators, corrected for edge effects by mirroring the sample \citep{Schuster+1985}.

@@

For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $11 \unit{M_{\odot}}$, in our fiducial simulations shown here. This is in stark contrast with observations from ground-based GW detectors, where heavy BHs with masses of about $30\unit{M_{\odot}}$ and higher have been common. There are two main reasons for this discrepancy. First, the population of BHs in the Milky Way (and, in particular, those detectable by LISA due to their recent formation times, see Figure~\ref{fig:fiducial_pdf_distributions}e) primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}. Secondly, ground-based detectors are strongly biased towards high mass systems, since they can be detected out to larger distances and thus a greater volume is probed. In contrast, LISA has no such bias and is, in principle, sensitive to inspiraling BHBHs throughout the entire Milky Way regardless of their mass, as long as we catch them when they are emitting in the LISA band. For this reason, LISA is more likely to detect the more common lower mass BHBHs.

@@

\begin{table}[tb]
    \centering
    \begin{tabular}{l|c|cccc}
        \hline
        \multirow{2}{*}{Component} & Formation & \multicolumn{4}{c}{Detectable} \\ \cline{2-6}
        & {\scriptsize All} & {\scriptsize All} & {\scriptsize BHBH} & {\scriptsize BHNS} & {\scriptsize NSNS} \\
        \hline
        Low-$\alpha$ disc & 42.5\% & 86\% & 89\% & 82\% & 85\%\\
        High-$\alpha$ disc & 42.5\% & 6\% & 5\% & 8\% & 12\%\\
        Bulge/bar & 15.0\% & 8\% & 6\% & 10\% & 3\%\\
    \end{tabular}
    \caption{Percentage of systems in each Galactic component.}
    \label{tab:component_fractions}
\end{table}
\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig4_1_galaxy_density_distribution.png}
    \includegraphics[width=\columnwidth]{fig4_2_detectable_distance_distribution_4yr.pdf}
    \caption{
    \textbf{Top:} A face-on view of the Galactic density distribution for detectable DCOs. We show the density distribution for the top 95\% of the sources, the rest are indicated by scatter points whose sizes correspond to their sampling weights. The coloured lines show the 75\% contour for each of the individual DCO types. The large cross passes through (0, 0) and helps to highlight the bias towards the position of the sun, which is indicated by the $\odot$.
    \textbf{Bottom:} As Figure~\ref{fig:fiducial_pdf_distributions}, but for the luminosity distance. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_1_galaxy_density_distribution.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/galaxy_creation_station.ipynb}{\faBook} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_2_detectable_distance_distribution_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:detectable_distance_dist}
\end{figure}

>>>>

\subsection{Properties of the detectable systems}\label{sec:fiducial_distributions}
In Fig.~\ref{fig:fiducial_pdf_distributions}, we show the 1D distribution of several individual parameters of the population of detectable binaries together with the 1- and 2-$\sigma$ uncertainties obtained via bootstrapping. These uncertainties represent the fluctuations in our results over different random instances of the Milky Way. The distributions shown here are approximated by kernel density estimators, corrected for edge effects by mirroring the sample \citep{Schuster+1985}.

@@

For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $11 \unit{M_{\odot}}$, in our fiducial simulations shown here. This is in stark contrast with observations from ground-based GW detectors, where heavy BHs with masses of about $30\unit{M_{\odot}}$ and higher have been common. There are two main reasons for this discrepancy. First, the population of BHs in the Milky Way (and, in particular, those detectable by LISA due to their recent formation times, see Fig.~\ref{fig:fiducial_pdf_distributions}e) primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}. Secondly, ground-based detectors are strongly biased towards high mass systems, since they can be detected out to larger distances and thus a greater volume is probed. In contrast, LISA has no such bias and is, in principle, sensitive to inspiraling BHBHs throughout the entire Milky Way regardless of their mass, as long as we catch them when they are emitting in the LISA band. For this reason, LISA is more likely to detect the more common lower mass BHBHs.

@@

\begin{table}[tb]
    \centering
    \begin{tabular}{l|c|cccc}
        \hline
        \multirow{2}{*}{Component} & Formation & \multicolumn{4}{c}{Detectable} \\ \cline{2-6}
        & {\scriptsize All} & {\scriptsize All} & {\scriptsize BHBH} & {\scriptsize BHNS} & {\scriptsize NSNS} \\
        \hline
        Low-$\alpha$ disc & 42.5\% & 86\% & 89\% & 82\% & 85\%\\
        High-$\alpha$ disc & 42.5\% & 6\% & 5\% & 8\% & 12\%\\
        Bulge/bar & 15.0\% & 8\% & 6\% & 10\% & 3\%\\
    \end{tabular}
    \caption{Percentage of systems in each Galactic component.}
    \label{tab:component_fractions}
\end{table}
\begin{figure}[tb]
    \centering
    \includegraphics[width=\columnwidth]{fig4_1_galaxy_density_distribution.png}
    \includegraphics[width=\columnwidth]{fig4_2_detectable_distance_distribution_4yr.pdf}
    \caption{
    \textbf{Top:} A face-on view of the Galactic density distribution for detectable DCOs. We show the density distribution for the top 95\% of the sources, the rest are indicated by scatter points whose sizes correspond to their sampling weights. The coloured lines show the 75\% contour for each of the individual DCO types. The large cross passes through (0, 0) and helps to highlight the bias towards the position of the sun, which is indicated by the $\odot$.
    \textbf{Bottom:} As Fig.~\ref{fig:fiducial_pdf_distributions}, but for the luminosity distance. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_1_galaxy_density_distribution.png}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/galaxy_creation_station.ipynb}{\faBook} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig4_2_detectable_distance_distribution_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:detectable_distance_dist}
\end{figure}

### 2021-11-19 paper/tex/sections/variations.tex
Change Figure->Fig.

\subsection{Detection rates}\label{sec:detection_rate_analysis}
In Figure~\ref{fig:detection_rates}, we show the expected number of LISA detections based on our simulations considering variations in the physical assumptions. We show the expected number of detections for BHBH, BHNS and NSNS systems in the top, middle and bottom panel respectively. All the rates and their uncertainties plotted in this figure are also provided in Table~\ref{tab:detection_rates}. 
In the sections that follow, we briefly explain the variations considered and discuss the most prominent trends.

@@

In Figure~\ref{fig:lower_mass_gap_variation}, we show the component mass distribution for all LISA detectable DCOs (BHBHs, BHNSs and NSNSs) for two different remnant mass prescriptions. The grey distribution uses the Fryer \textit{delayed} remnant mass prescription \citep{Fryer+2012}, which is our fiducial assumption (model \modFid{}). This prescription produces compact objects in the lower mass gap ($2.5 \unit{M_{\odot}} \le m \le 5 \unit{M_{\odot}}$) and indeed we find that, of the LISA detectable DCOs, approximately \BHBHatLeastOneLowerMassGapPerc{}\% of BHBHs, \BHNSatLeastOneLowerMassGapPerc{}\% of BHNSs and \NSNSatLeastOneLowerMassGapPerc{}\% of NSNSs have at least one component in the lower mass gap. Overall, weighting by the relative detection rates, this gives that 55\% of our predicted LISA DCO detections would have at least one component in the lower mass gap when using this remnant mass prescription.

@@

In Figure~\ref{fig:ecc_low_kick_variation}, we investigate how decreasing the magnitude of natal kicks from core-collapse supernovae affects the eccentricity distribution of LISA detectable DCOs. For reference, we show the mean fiducial distributions (model \modFid{}) as dashed lines (see Fig.~\ref{fig:fiducial_pdf_distributions}d for full comparison). In the main curves, we reduce the velocity dispersion for core-collapse supernovae from $265 \unit{km}{s^{-1}}$ to $30 \unit{km}{s^{-1}}$ (model \modSigLower{}).

>>>>

\subsection{Detection rates}\label{sec:detection_rate_analysis}
In Fig.~\ref{fig:detection_rates}, we show the expected number of LISA detections based on our simulations considering variations in the physical assumptions. We show the expected number of detections for BHBH, BHNS and NSNS systems in the top, middle and bottom panel respectively. All the rates and their uncertainties plotted in this figure are also provided in Table~\ref{tab:detection_rates}. 
In the sections that follow, we briefly explain the variations considered and discuss the most prominent trends.

@@

In Fig.~\ref{fig:lower_mass_gap_variation}, we show the component mass distribution for all LISA detectable DCOs (BHBHs, BHNSs and NSNSs) for two different remnant mass prescriptions. The grey distribution uses the Fryer \textit{delayed} remnant mass prescription \citep{Fryer+2012}, which is our fiducial assumption (model \modFid{}). This prescription produces compact objects in the lower mass gap ($2.5 \unit{M_{\odot}} \le m \le 5 \unit{M_{\odot}}$) and indeed we find that, of the LISA detectable DCOs, approximately \BHBHatLeastOneLowerMassGapPerc{}\% of BHBHs, \BHNSatLeastOneLowerMassGapPerc{}\% of BHNSs and \NSNSatLeastOneLowerMassGapPerc{}\% of NSNSs have at least one component in the lower mass gap. Overall, weighting by the relative detection rates, this gives that 55\% of our predicted LISA DCO detections would have at least one component in the lower mass gap when using this remnant mass prescription.

@@

In Fig.~\ref{fig:ecc_low_kick_variation}, we investigate how decreasing the magnitude of natal kicks from core-collapse supernovae affects the eccentricity distribution of LISA detectable DCOs. For reference, we show the mean fiducial distributions (model \modFid{}) as dashed lines (see Fig.~\ref{fig:fiducial_pdf_distributions}d for full comparison). In the main curves, we reduce the velocity dispersion for core-collapse supernovae from $265 \unit{km}{s^{-1}}$ to $30 \unit{km}{s^{-1}}$ (model \modSigLower{}).
